---
title: "Homework 1"
author: "Hyder Said"
date: "2026-09-21"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Problem 1

```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

The penguins dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables. Each row represents an individual penguin. Important variables include `species` (`r paste(levels(penguins$species), collapse = ", ")`), `island` (`r paste(levels(penguins$island), collapse = ", ")`), `sex`, and `year`, as well as measurements of bill length, bill depth, flipper length, and body mass. The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)` mm.

```{r penguin_scatterplot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguin_scatterplot.png")
```


# Problem 2
```{r create_df}
set.seed(0)

p2_df = tibble(
  num_var = rnorm(n = 10),
  log_var = num_var > 0,
  char_var = c(
    "one", "two", "three", "four", "five",
    "six", "seven", "eight", "nine", "ten"
  ),
  factor_var = factor(
    c("A", "B", "C", "A", "B", "C", "A", "B", "C", "A")
  )
)

p2_df
```


```{r calculate_mean}
mean(pull(p2_df, num_var))
mean(pull(p2_df, log_var))
mean(pull(p2_df, char_var))
mean(pull(p2_df, factor_var))
```

I used `pull()` to try to calculate the mean of each variable. The mean worked for `num_var`. It also worked for `log_var` because R treats `TRUE` as 1 and `FALSE` as 0, so the result is the proportion of sampled values greater than zero. The means of `char_var` and `factor_var` did not work and returned missing values.

```{r convert_variables, results = "hide"}
as.numeric(pull(p2_df, log_var))
as.numeric(pull(p2_df, char_var))
as.numeric(pull(p2_df, factor_var))
```

After converting the variables to numeric, `log_var` became 1 for `TRUE` and 0 for `FALSE`, which explains why its mean worked. Converting `char_var` produced missing values because R can't interpret words like `"one"` and `"two"` as numbers. Converting `factor_var` produced the codes 1, 2, and 3 for its levels, but these are just category labels, so taking their mean doesn't make much sense.



